---
title: "WSLS analysis AmydACC"
output: html_notebook
---

### Import dependencies
```{r}
library(emmeans)
```

### Import the dataframe
```{r}
WSLS_csv_location <- "/Volumes/project/3025011.02/pre-processed/behaviour/combined_dataframe_for_r.csv"

WSLS_dataframe <- read.csv(WSLS_csv_location, header = TRUE, sep = ";")
```

# Win-stay analysis
### Run a very simple regression model
```{r}
# Run the raw model
model <- lm(WS ~ subjectively_correct_one_back_int + session, data = WSLS_dataframe)

summary(model)
```

### Now code dummy variables
```{r}
# Code dummy variables for the stimulation conditions (now sessions)
WSLS_dataframe$session <- factor(WSLS_dataframe$session, levels = c(2, 3, 4))
WSLS_dataframe$congruency <- factor(WSLS_dataframe$congruency, levels = c(0, 1))
WSLS_dataframe$volatility <- factor(WSLS_dataframe$volatility, levels = c(0, 1))
WSLS_dataframe$subjectively_correct_one_back_int <- factor(WSLS_dataframe$subjectively_correct_one_back_int, levels = c(0 ,1))
contrasts(WSLS_dataframe$session) <- contr.treatment(nlevels(WSLS_dataframe$session), base = 3) # this means that 3 is the baseline group
WSLS_dataframe$session
contrasts(WSLS_dataframe$session)

session_2 <- c(1, 0, 0)
session_3 <- c(0, 1, 0)
session_4 <- c(0, 0, 1)

contrasts(WSLS_dataframe$session) <- cbind(session_2, session_3, session_2)
```

### Now run the model with the dummy variables
This might require logistic regression instead
```{r}
model <- lm(WS ~ subjectively_correct_one_back_int + session, data = WSLS_dataframe)
summary(model)
```

### You can also do a pairwise comparison instead of contrasts
```{r}
session_emm <- emmeans(model, ~ session, type = "response")
pairwise_comp <- pairs(session_emm)
print(pairwise_comp)
```

#### Make different models that also test the influence of volatility or congruency
```{r}
model <- lm(WS ~ congruency + session, data = WSLS_dataframe)
summary(model)
```
```{r}
model <- lm(WS ~ congruency + volatility + session, data = WSLS_dataframe)
summary(model)
```

# Lose-shift analysis
### Run a very simple regression model
```{r}
model <- lm(LS ~ subjectively_correct_one_back_int + same_choice_as_one_back, data = WSLS_dataframe)

summary(model)
```
