---
title: "WSLS analysis AmydACC"
output: html_notebook
---

### Import dependencies

```{r}
library(car)
library(mlogit)
library(glue)
library(lmtest)
library(lme4)
```

### Import the dataframe

```{r}
WSLS_csv_location <- "/Volumes/project/3025011.02/pre-processed/behaviour/combined_dataframe_for_r.csv"

WSLS_dataframe <- read.csv(WSLS_csv_location, header = TRUE, sep = ";")
```

### Recode categorical variables as factors

```{r}
WSLS_dataframe$response <- as.factor(WSLS_dataframe$response)
is.factor(WSLS_dataframe$response)
WSLS_dataframe$WS_subjective <- as.factor(WSLS_dataframe$WS_subjective)
is.factor(WSLS_dataframe$WS_subjective)
WSLS_dataframe$LS_subjective <- as.factor(WSLS_dataframe$LS_subjective)
is.factor(WSLS_dataframe$LS_subjective)
WSLS_dataframe$WS_objective <- as.factor(WSLS_dataframe$WS_objective)
is.factor(WSLS_dataframe$WS_objective)
WSLS_dataframe$LS_objective <- as.factor(WSLS_dataframe$LS_objective)
is.factor(WSLS_dataframe$LS_objective)
WSLS_dataframe$choicestickiness <- as.factor(WSLS_dataframe$choicestickiness)
is.factor(WSLS_dataframe$choicestickiness)
WSLS_dataframe$volatility <- as.factor(WSLS_dataframe$volatility)
is.factor(WSLS_dataframe$volatility)
WSLS_dataframe$subject_id <- as.factor(WSLS_dataframe$subject_id)
is.factor(WSLS_dataframe$subject_id)

WSLS_dataframe$RT_s <- as.numeric(WSLS_dataframe$RT_s)
is.numeric(WSLS_dataframe$RT_s)
WSLS_dataframe$response_bias <- as.numeric(WSLS_dataframe$response_bias)
is.numeric(WSLS_dataframe$response_bias)

head(WSLS_dataframe)
```

### Relevel the data

This ensures that the baseline of any factor is set to something like session 1 or no TUS

```{r}
#WSLS_dataframe$volatility <- relevel(WSLS_dataframe$volatility, 'stable')
```

# Multinomial logistic regression

### Put data in long format

```{r}
WS_subjective_dataframe_long <- mlogit.data(WSLS_dataframe, choice = 'WS_subjective', shape = 'wide')
print(WS_subjective_dataframe_long)
LS_subjective_dataframe_long <- mlogit.data(WSLS_dataframe, choice = 'LS_subjective', shape = 'wide')
print(LS_subjective_dataframe_long)
WS_objective_dataframe_long <- mlogit.data(WSLS_dataframe, choice = 'WS_objective', shape = 'wide')
print(WS_subjective_dataframe_long)
LS_objective_dataframe_long <- mlogit.data(WSLS_dataframe, choice = 'LS_objective', shape = 'wide')
print(LS_subjective_dataframe_long)
```

# Subjective Win-Stay models

### Multinomial model

```{r}
multinomail_model_1 <- mlogit(WS_subjective ~ 1, data = WS_subjective_dataframe_long, reflevel = '1', method = 'nr', print.level = 0)

multinomail_model_2 <- mlogit(WS_subjective ~ 1 | response_bias, data = WS_subjective_dataframe_long, reflevel = '1', method = 'nr', print.level = 0)

multinomail_model_3 <- mlogit(WS_subjective ~ 1 | response_bias + time_since_switch, data = WS_subjective_dataframe_long, reflevel = '1', method = 'nr', print.level = 0)

multinomail_model_4 <- mlogit(WS_subjective ~ 1 | response_bias + time_since_switch + choicestickiness, data = WS_subjective_dataframe_long, reflevel = '1', method = 'nr', print.level = 0)

multinomail_model_5 <- mlogit(WS_subjective ~ 1 | response_bias + time_since_switch + choicestickiness + volatility, data = WS_subjective_dataframe_long, reflevel = '1', method = 'nr', print.level = 0)

multinomail_model_6 <- mlogit(WS_subjective ~ 1 | response_bias + time_since_switch + choicestickiness + volatility + choicestickiness:volatility, data = WS_subjective_dataframe_long, reflevel = '1', method = 'nr', print.level = 0)

multinomail_model_7 <- mlogit(WS_subjective ~ 1 | response_bias + time_since_switch + choicestickiness + volatility + choicestickiness:volatility + congruency, data = WS_subjective_dataframe_long, reflevel = '1', method = 'nr', print.level = 0)

multinomail_model_8 <- mlogit(WS_subjective ~ 1 | response_bias + time_since_switch + choicestickiness + volatility + choicestickiness:volatility + congruency + congruency:volatility, data = WS_subjective_dataframe_long, reflevel = '1', method = 'nr', print.level = 0)

multinomail_model_9 <- mlogit(WS_subjective ~ 1 | response_bias + time_since_switch + choicestickiness + volatility + choicestickiness:volatility + congruency + congruency:volatility + choicestickiness:congruency, data = WS_subjective_dataframe_long, reflevel = '1', method = 'nr', print.level = 0)

multinomail_model_10 <- mlogit(WS_subjective ~ 1 | response_bias + time_since_switch + choicestickiness + volatility + choicestickiness:volatility + congruency + congruency:volatility + choicestickiness:congruency + time_since_switch:congruency, data = WS_subjective_dataframe_long, reflevel = '1', method = 'nr', print.level = 0)

multinomail_model_11 <- mlogit(WS_subjective ~ 1 | response_bias + time_since_switch + choicestickiness + volatility + choicestickiness:volatility + congruency + time_since_switch:volatility, data = WS_subjective_dataframe_long, reflevel = '1', method = 'nr', print.level = 0)

glmer_model <- glmer(WS_subjective ~ response_bias + time_since_switch + choicestickiness + volatility + choicestickiness:volatility + congruency + time_since_switch:volatility + (1 | subject_id), data = WSLS_dataframe, family = binomial) # use this instead, included random intercept
```

```{r}
summary(multinomail_model_1)
```

```{r}
summary(multinomail_model_2)
```

```{r}
summary(multinomail_model_3)
```

```{r}
summary(multinomail_model_4)
```

```{r}
summary(multinomail_model_5)
```

```{r}
summary(multinomail_model_6)
```

```{r}
summary(multinomail_model_7)
```

```{r}
summary(multinomail_model_8)
```

```{r}
summary(multinomail_model_9)
```

```{r}
summary(multinomail_model_10)
```

```{r}
summary(multinomail_model_11)
```

### Look at the coefficients of the model

```{r}
exp(multinomail_model_3$coefficients)
```

### Look at the confidence intervals for the coefficients of the model

Crossing 1 is a bad sign

```{r}
exp(confint(multinomail_model_3))
```

### Model comparison

```{r}
AIC(multinomail_model_1, multinomail_model_2, multinomail_model_3, multinomail_model_4, multinomail_model_5, multinomail_model_6, multinomail_model_7, multinomail_model_8, multinomail_model_9, multinomail_model_10, multinomail_model_11)

lrtest(multinomail_model_1, multinomail_model_2)

lrtest(multinomail_model_2, multinomail_model_3)

lrtest(multinomail_model_3, multinomail_model_4)

lrtest(multinomail_model_4, multinomail_model_5)
```

```{r}
lrtest(multinomail_model_5, multinomail_model_6)

lrtest(multinomail_model_6, multinomail_model_7)

lrtest(multinomail_model_7, multinomail_model_8)

lrtest(multinomail_model_8, multinomail_model_9)
```

```{r}
lrtest(multinomail_model_5, multinomail_model_6)

lrtest(multinomail_model_6, multinomail_model_7)

lrtest(multinomail_model_7, multinomail_model_8)

lrtest(multinomail_model_9, multinomail_model_10)

lrtest(multinomail_model_10, multinomail_model_11)
```

# Subjective lose-shift

### Multinomial model

```{r}
multinomail_model_1 <- mlogit(LS_subjective ~ 1, data = LS_subjective_dataframe_long, reflevel = '1', method = 'nr', print.level = 0)

multinomail_model_2 <- mlogit(LS_subjective ~ 1 | response_bias, data = LS_subjective_dataframe_long, reflevel = '1', method = 'nr', print.level = 0)

multinomail_model_3 <- mlogit(LS_subjective ~ 1 | response_bias + choicestickiness, data = LS_subjective_dataframe_long, reflevel = '1', method = 'nr', print.level = 0)

multinomail_model_4 <- mlogit(LS_subjective ~ 1 | response_bias + choicestickiness + volatility, data = LS_subjective_dataframe_long, reflevel = '1', method = 'nr', print.level = 0)

multinomail_model_5 <- mlogit(LS_subjective ~ 1 | response_bias + choicestickiness + volatility + choicestickiness:volatility, data = LS_subjective_dataframe_long, reflevel = '1', method = 'nr', print.level = 0)

multinomail_model_6 <- mlogit(LS_subjective ~ 1 | response_bias + choicestickiness + volatility + choicestickiness:volatility + congruency, data = LS_subjective_dataframe_long, reflevel = '1', method = 'nr', print.level = 0)

multinomail_model_7 <- mlogit(LS_subjective ~ 1 | response_bias + choicestickiness + volatility + choicestickiness:volatility + congruency + congruency:volatility, data = LS_subjective_dataframe_long, reflevel = '1', method = 'nr', print.level = 0)

multinomail_model_8 <- mlogit(LS_subjective ~ 1 | response_bias + choicestickiness + volatility + choicestickiness:volatility + congruency + congruency:volatility + time_since_switch, data = LS_subjective_dataframe_long, reflevel = '1', method = 'nr', print.level = 0)

multinomail_model_9 <- mlogit(LS_subjective ~ 1 | response_bias + choicestickiness + volatility + choicestickiness:volatility + congruency + congruency:volatility + time_since_switch + time_since_switch:volatility, data = LS_subjective_dataframe_long, reflevel = '1', method = 'nr', print.level = 0)

AIC(multinomail_model_1, multinomail_model_2, multinomail_model_3, multinomail_model_4, multinomail_model_5, multinomail_model_6, multinomail_model_7, multinomail_model_8, multinomail_model_9)
```

```{r}
summary(multinomail_model_1)
```

```{r}
summary(multinomail_model_2)
```

```{r}
summary(multinomail_model_3)
```

```{r}
summary(multinomail_model_4)
```

```{r}
summary(multinomail_model_5)
```

```{r}
summary(multinomail_model_6)
```

```{r}
summary(multinomail_model_7)
```

```{r}
summary(multinomail_model_8)
```

```{r}
summary(multinomail_model_9)
```

### Look at the coefficients of the model

```{r}
exp(multinomail_model_3$coefficients)
```

### Look at the confidence intervals for the coefficients of the model

Crossing 1 is a bad sign

```{r}
exp(confint(multinomail_model_3))
```

### Model comparison

```{r}
AIC(multinomail_model_1, multinomail_model_2, multinomail_model_3, multinomail_model_4, multinomail_model_5, multinomail_model_6, multinomail_model_7, multinomail_model_8, multinomail_model_9)

lrtest(multinomail_model_1, multinomail_model_2)

lrtest(multinomail_model_2, multinomail_model_3)

lrtest(multinomail_model_3, multinomail_model_4)
```

```{r}
lrtest(multinomail_model_4, multinomail_model_5)

lrtest(multinomail_model_5, multinomail_model_6)

lrtest(multinomail_model_6, multinomail_model_7)

lrtest(multinomail_model_7, multinomail_model_8)
```

# Objective Win-Stay models

### Multinomial model

```{r}
multinomail_model_1 <- mlogit(WS_objective ~ 1, data = WS_objective_dataframe_long, reflevel = '1', method = 'nr', print.level = 0)

multinomail_model_2 <- mlogit(WS_objective ~ 1 | response_bias, data = WS_objective_dataframe_long, reflevel = '1', method = 'nr', print.level = 0)

multinomail_model_3 <- mlogit(WS_objective ~ 1 | response_bias + choicestickiness, data = WS_objective_dataframe_long, reflevel = '1', method = 'nr', print.level = 0)

multinomail_model_4 <- mlogit(WS_objective ~ 1 | response_bias + choicestickiness + volatility, data = WS_objective_dataframe_long, reflevel = '1', method = 'nr', print.level = 0)

multinomail_model_5 <- mlogit(WS_objective ~ 1 | response_bias + choicestickiness + volatility + choicestickiness:volatility, data = WS_objective_dataframe_long, reflevel = '1', method = 'nr', print.level = 0)

multinomail_model_6 <- mlogit(WS_objective ~ 1 | response_bias + choicestickiness + volatility + choicestickiness:volatility + congruency, data = WS_objective_dataframe_long, reflevel = '1', method = 'nr', print.level = 0)

multinomail_model_7 <- mlogit(WS_objective ~ 1 | response_bias + choicestickiness + volatility + choicestickiness:volatility + congruency + time_since_switch, data = WS_objective_dataframe_long, reflevel = '1', method = 'nr', print.level = 0)

multinomail_model_8 <- mlogit(WS_objective ~ 1 | response_bias + choicestickiness + volatility + choicestickiness:volatility + congruency + time_since_switch + time_since_switch:volatility, data = WS_objective_dataframe_long, reflevel = '1', method = 'nr', print.level = 0)

AIC(multinomail_model_1, multinomail_model_2, multinomail_model_3, multinomail_model_4, multinomail_model_5, multinomail_model_6, multinomail_model_7, multinomail_model_8)
```
# Win-stay analysis

### Run a very simple regression model

```{r}
model1 <- glm(WS_subjective ~ choicestickiness, data = WS_subjective_dataframe, family = binomial())

summary(model1)
```

### Add the training as the factor 'session

```{r}
model2 <- glm(WS_subjective ~ choicestickiness + volatility, data = WS_subjective_dataframe, family = binomial())

summary(model2)
```

### Compare factors

```{r}
model_chisquare <- model1$null.deviance - model1$deviance
model_df <- model1$df.null - model1$df.residual
model_chisquare_p <- 1 - pchisq(model_chisquare, model_df)
glue('Model Chi-square test: x^2({model_df}) = {model_chisquare}, p = {model_chisquare_p}')
```

### Calculate different R\^2's

```{r}
model_R1HL <- model_chisquare / model1$null.deviance
model_R1CS <- 1 - exp( -(model1$null.deviance - model1$deviance) / length(model1$fitted.values))
model_R1N <- model_R1CS / (1 -(exp(-(model1$null.deviance / length(model1$fitted.values)))))
glue('Hosmer and Lemeshows pseudo-R^2 = {model_R1HL}')
glue('Cox and Snells pseudo-R^2 = {model_R1CS}')
glue('Nagelkerke pseudo-R^2 = {model_R1N}')
```

## Odd-ratio

Change in odds

```{r}
glue('If values are greater than 1, there is a positive correlation between the predictor and outcome')
glue('If lower than 1, there is a negative correlation between the predictor and the outcome')
glue('Odds ratio for factors is:')
exp(model1$coefficients)
glue('People who stick with their choice when going from n-2 to n-1 are 1.83 times as likely to stick with their choice again')
exp(confint(model1))
```

### Check assumptions

```{r}
WS_subjective_dataframe$predicted.probabilities <- fitted(model1)
WS_subjective_dataframe$standardized.residuals <- rstandard(model1)
WS_subjective_dataframe$studentized.residuals <- rstudent(model1)
WS_subjective_dataframe$dfbeta <- dfbeta(model1)
WS_subjective_dataframe$dffit <- dffits(model1)
WS_subjective_dataframe$leverage <- hatvalues(model1)

WS_subjective_dataframe[, c('WS_subjective', 'choicestickiness', 'volatility', 'predicted.probabilities')]
```

```{r}
WS_subjective_dataframe[, c('leverage', 'studentized.residuals', 'dfbeta')]
```

### Testing collinearity assumption

```{r}
print('VIF values:')
print('VIF over 10 is a sign of collinearity')
vif(model2)
print('Tolerance')
1/vif(model2)
```

### Testing log linnearity assumption

```{r}
print('Check the linearity to the log of the outcome variable for every factor:')
WS_subjective_dataframe$logchoicestickiness <- log(WS_subjective_dataframe$choicestickiness)*WS_subjective_dataframe$choicestickiness
WS_subjective_dataframe$logvolatility <- log(WS_subjective_dataframe$volatility)*WS_subjective_dataframe$volatility
head(WS_subjective_dataframe)

print('Then create a model with both he normal factors and the log of those factors')
print('If any of the log factors are significant, that means that the assumption has been violated')
model_log_linnearity <- glm(WS_subjective ~ choicestickiness + volatility + logchoicestickiness + logvolatility, data = WS_subjective_dataframe, family = binomial())
summary(model_log_linnearity)
```

### Now code dummy variables

```{r}
# Code dummy variables for the stimulation conditions (now sessions)
WSLS_dataframe$session <- factor(WSLS_dataframe$session, levels = c(2, 3, 4))
WSLS_dataframe$congruency <- factor(WSLS_dataframe$congruency, levels = c(0, 1))
WSLS_dataframe$volatility <- factor(WSLS_dataframe$volatility, levels = c(0, 1))
WSLS_dataframe$subjectively_correct_one_back_int <- factor(WSLS_dataframe$subjectively_correct_one_back_int, levels = c(0 ,1))
contrasts(WSLS_dataframe$session) <- contr.treatment(nlevels(WSLS_dataframe$session), base = 3) # this means that 3 is the baseline group
WSLS_dataframe$session
contrasts(WSLS_dataframe$session)

session_2 <- c(1, 0, 0)
session_3 <- c(0, 1, 0)
session_4 <- c(0, 0, 1)

contrasts(WSLS_dataframe$session) <- cbind(session_2, session_3, session_2)
```
