---
title: "Drift-diffusion model AmydACC"
output: html_notebook
---

# Load dependancies

```{r}
rm(list=ls()) # clear workspace

set.seed(1) # set random seed so results will replicate
# Note:  Seed will be reset in each section below, to re-initialize before running  
# chunks of code, so that results will replicate even if portions of code are 
# skipped or re-run on a particular session.

options(digits = 3) # by default, print results to three decimal digits
# Below packages should be loaded into workspace; each (and any dependencies)
# must be installed from CRAN before loading via library().
library(rtdists) # DDM functions rdiffusion(), ddiffusion(), qdiffusion()
library(RWiener) # alternate DDM functions rwiener(), deviance.wdm(), logLik.wdm()
library(tidyr)   # unnest() function
library(dplyr)   # mutate() and arrange() functions
library(purrr)   # split() and map() functions
library(ggplot2)   # graphing
library(gridExtra) # for multiple ggplots in one window
library(gt)        # grammar of tables
library(xml2)      # save formatted tables as rtf 
library(reshape2)  # for melt() function

setwd("/Users/kenneth.van.der.zee/Library/Mobile Documents/com~apple~CloudDocs/Werk/Courses/Drift diffusion modelling")
data.pathname = "/Users/kenneth.van.der.zee/Library/Mobile Documents/com~apple~CloudDocs/Werk/Courses/Drift diffusion modelling/data/"

# Then load the DMC routines:
source("dmc/dmc.R")

set.seed(1) # reset random seed so results for each portion of code will replicate
```

# Load some data in
```{r}
combined_results_dataframe <- read.csv("/Volumes/affneu/kenvdzee/Documents/phd_1_analysis/pre-processed/data_for_modelling/combined_dataframe_for_r.csv", header=TRUE, sep=";")

print(combined_results_dataframe)
```
# Take one participant and one session for the first excercises
```{r}
str(combined_results_dataframe[c("subject_id","session")])
combined_results_dataframe_subset <- combined_results_dataframe[
  combined_results_dataframe$subject_id == 5 &
  combined_results_dataframe$session    == 4,
]
print(combined_results_dataframe_subset)
```
# Combine the two conditions you want to analyse
```{r}
combined_results_dataframe_subset <- mutate(combined_results_dataframe_subset, congruency_performance = case_when(
  (congruency=="1" & objectively_correct_int==1) ~ "congruent_correct",
  (congruency=="1" & objectively_correct_int==0) ~ "congruent_incorrect",
  (congruency=="0" & objectively_correct_int==1) ~ "incongruent_correct",
  (congruency=="0" & objectively_correct_int==0) ~ "incongruent_incorrect" ))

combined_results_dataframe_subset_table <- table(combined_results_dataframe_subset$congruency, combined_results_dataframe_subset$objectively_correct_int)
combined_results_dataframe_subset_table

(observed_prop_correct.congruent <- combined_results_dataframe_subset_table[1,2]/(combined_results_dataframe_subset_table[1,2]+combined_results_dataframe_subset_table[2,2]))
(observed_prop_correct.incongruent <- combined_results_dataframe_subset_table[1,1]/(combined_results_dataframe_subset_table[1,1]+combined_results_dataframe_subset_table[2,1]))
```

# This is the demo data
```{r}
file1 <- paste(data.pathname, "testdata.csv", sep="")
data1 <- read.csv(file1, header=TRUE)   # 500 observations of 3 variables

# We'll add a variable to code 4 types of trial based on configuration 
# of stimulus and response (s1.r1, s1.r2, s2.r1, s2.r2)
data1 <- mutate(data1, myType = case_when(
  (S=="s1" & R=="r1") ~ "s1.r1",
  (S=="s1" & R=="r2") ~ "s1.r2",
  (S=="s2" & R=="r1") ~ "s2.r1",
  (S=="s2" & R=="r2") ~ "s2.r2" ))

temp1 <- table(data1$S, data1$R) 
temp1
#     r1  r2
# s1 114  36   
# s2  61 289  
(observed_prop_correct.s1 <- temp1[1,1]/(temp1[1,1]+temp1[1,2]))
(observed_prop_correct.s2 <- temp1[2,2]/(temp1[2,1]+temp1[2,2]))
# I.e., s1: 114 correct out of 150 s1 trials (76% correct); 
# and s2: 289 correct out of 350 s2 trials (82.6% correct).
```

```{r}
temp2 <- list(100*observed_prop_correct.congruent, 100*observed_prop_correct.incongruent)
names(temp2) <- c("Congruent", "Incongruent")
barplot(unlist(temp2), ylim=c(0,100), ylab="Percent Correct", xlab="Congruency")
```
```{r}
# Also a quick look at mean/SD of correct and incorrect RTs in the dataset
temp3 <- aggregate(combined_results_dataframe_subset$RT_s, list(combined_results_dataframe_subset$congruency_performance), mean)
temp4 <- aggregate(combined_results_dataframe_subset$RT_s, list(combined_results_dataframe_subset$congruency_performance), sd)

temp3
temp4
```
```{r}
# Quick frequency histograms of RTs for each stimulus, broken down
# by correct and incorrect responses
congruent_correct <- combined_results_dataframe_subset[combined_results_dataframe_subset$congruency_performance=="congruent_correct",]
congruent_incorrect <- combined_results_dataframe_subset[combined_results_dataframe_subset$congruency_performance=="congruent_incorrect",]
incongruent_correct <- combined_results_dataframe_subset[combined_results_dataframe_subset$congruency_performance=="incongruent_correct",]
incongruent_incorrect <- combined_results_dataframe_subset[combined_results_dataframe_subset$congruency_performance=="incongruent_incorrect",]

par(mfrow=c(2,2))
hist(congruent_correct$RT_s, prob=FALSE, main="congruent_correct", breaks=seq(0,2.5,by=0.1),
     xlab="RT (sec)", ylim=c(0,125))
hist(congruent_incorrect$RT_s, prob=FALSE, main="congruent_incorrect", breaks=seq(0,2.5,by=0.1),
     xlab="RT (sec)", ylim=c(0,125))
hist(incongruent_correct$RT_s, prob=FALSE, main="incongruent_correct", breaks=seq(0,2.5,by=0.1),
     xlab="RT (sec)", ylim=c(0,125))
hist(incongruent_incorrect$RT_s, prob=FALSE, main="incongruent_incorrect", breaks=seq(0,2.5,by=0.1), 
     xlab="RT (sec)", ylim=c(0,125))
```

```{r}
###### Also find RT quantiles (.1, .3, .5, .7, .9) for each S, R
my.quantiles <- c(0.1, 0.3, 0.5, 0.7, 0.9)
quantiles.combined_results_dataframe_subset <- combined_results_dataframe_subset %>% group_by(congruency, objectively_correct_int) %>% nest() %>%
  mutate(quantiles=map(data,
                       ~ as.data.frame(t(quantile(.x$RT_s, probs=my.quantiles))))) %>%
  unnest(quantiles) %>%
  gather("quantile", "RT_s", '10%':'90%') %>%
  arrange(congruency, objectively_correct_int)

###### Approximate CDFs based on quantile RTs; we'll use these later ######
data1_q_s1.r1 <- quantile(combined_results_dataframe_subset[combined_results_dataframe_subset$congruency_performance=="congruent_correct", "RT_s"], probs=my.quantiles)
data1_q_s1.r2 <- quantile(combined_results_dataframe_subset[combined_results_dataframe_subset$congruency_performance=="congruent_incorrect", "RT_s"], probs=my.quantiles)
data1_q_s2.r1 <- quantile(combined_results_dataframe_subset[combined_results_dataframe_subset$congruency_performance=="incongruent_correct", "RT_s"], probs=my.quantiles)
data1_q_s2.r2 <- quantile(combined_results_dataframe_subset[combined_results_dataframe_subset$congruency_performance=="incongruent_incorrect", "RT_s"], probs=my.quantiles)



####### Given some "reasonable" starting values, run the DDM on our test dataset 

# Here, we'll use rwiener() to generate 150 s1 trials and 350 s2 trials.
# Note conventions: tau is Ter, beta is z, and delta>1 indicates that
# drift rate slopes in the correct direction for that stimuli.
# It returns a vector of resp="upper"/"lower" where "upper" is defined
# as r2 and q=RT in sec.
set.seed(0) 
dat.congruent <- rwiener(n=334, alpha=1.0, tau=0.2, beta=0.5, delta= -1)
dat.incongruent <- rwiener(n=334, alpha=1.0, tau=0.2, beta=0.5, delta=1.25)
dat.congruent$congruency <- rep("congruent", 334)
dat.incongruent$congruency <- rep("incongruent", 334)
sim.data <- rbind(dat.congruent, dat.incongruent)
sim.data
# rename "q" to "RT" for clarity
names(sim.data)[names(sim.data) == 'q'] <- 'RT'
# change resp="lower"/"upper" to r="r1"/"r2"
sim.data <- mutate(sim.data, objectively_correct_int = case_when(
  (resp=="correct") ~ "1",
  (resp=="incorrect") ~ "0" ))
# what happened on first few trials?
head(sim.data)
#      RT  resp  S  R
# 1 0.382 lower s1 r1
# 2 0.495 lower s1 r1
# 3 0.712 lower s1 r1
# 4 0.269 upper s1 r2
# 5 0.328 lower s1 r1
# 6 0.378 lower s1 r1

# check the results: s1 should (usually) map to r1 and s2 to r2
(x <- table(sim.data$congruency, sim.data$objectively_correct_int))
x
#     r1  r2
# s1 116  34
# s2  88 262
(pred_prop_correct_DDM.s1 <- x[1,1]/(x[1,1]+x[1,2]))   # accuracy to s1 = 0.77
(pred_prop_correct_DDM.s2 <- x[2,2]/(x[2,1]+x[2,2]))   # accuracy to s2 = 0.75
```

###### Manuscript FIGURE 5A: accuracy of predicted data
par(mfrow=c(1,1))
temp3 <- list(100*pred_prop_correct_DDM.s1, 100*pred_prop_correct_DDM.s2)
names(temp3) <- c("s1", "s2")
(a<-barplot(unlist(temp3), ylim=c(0,100), ylab="Percent Correct", xlab="Stimulus",
        main="PREDICTED Accuracy"))
points(x=a, unlist(temp2), col="red")  # overlay dots representing empirical accuracy

###### MANUSCRIPT FIGURE 5B: frequency histograms of predicted RT ######
P_s1r1 <- sim.data[sim.data$S=="s1"&sim.data$R=="r1",]
P_s2r2 <- sim.data[sim.data$S=="s2"&sim.data$R=="r2",]
P_s1r2 <- sim.data[sim.data$S=="s1"&sim.data$R=="r2",]
P_s2r1 <- sim.data[sim.data$S=="s2"&sim.data$R=="r1",]
par(mfrow=c(2,2))
hist(P_s1r1$RT, prob=FALSE, main="PREDICTED: Stimulus s1, Correct", breaks=seq(0,2.5,by=0.1),
     xlab="RT (sec)", ylim=c(0,80))
hist(P_s2r2$RT, prob=FALSE, main="PREDICTED: Stimulus s2, Correct", breaks=seq(0,2.5,by=0.1),
     xlab="RT (sec)", ylim=c(0,80))
hist(P_s1r2$RT, prob=FALSE, main="PREDICTED: Stimulus s1, Error", breaks=seq(0,2.5,by=0.1),
     xlab="RT (sec)", ylim=c(0,80))
hist(P_s2r1$RT, prob=FALSE, main="PREDICTED: Stimulus s2, Error", breaks=seq(0,2.5,by=0.1), 
     xlab="RT (sec)", ylim=c(0,80))




###### Manuscript FIGURE 5C: Compare predicted quantiles vs. observed data ######
# This figure is used to illustrate the chi-square method of parameter estimation,
# but for illustration, we'll just use our "sensible" DDM parameters 

# We already calculated above quantiles for the observed data: 
#        data1_q_s1.r1, data1_q_s1.r2, data1_q_s2.r1, data1_q_s2.r2
# and predicted proportion correct in observed data:
#        observed_prop_correct.s1 and observed_prop_correct.s2
# Next compute quantiles for the DDM; NB for simplicity, we'll use the 
# predicted data obtained above with our "first guess" of parameter values;
# we already have accuracy for the predicted data
#        pred_prop_correct_DDM.s1, pred_prop_correct_DDM.s2

# Next generate DDM quantiles for the predicted data
my.quantiles <- c(0.1, 0.3, 0.5, 0.7, 0.9)
pred_q_s1.r1 <- quantile(sim.data[sim.data$S=="s1"&sim.data$R=="r1", "RT"], probs=my.quantiles)
pred_q_s1.r2 <- quantile(sim.data[sim.data$S=="s1"&sim.data$R=="r2", "RT"], probs=my.quantiles)
pred_q_s2.r1 <- quantile(sim.data[sim.data$S=="s2"&sim.data$R=="r1", "RT"], probs=my.quantiles)
pred_q_s2.r2 <- quantile(sim.data[sim.data$S=="s2"&sim.data$R=="r2", "RT"], probs=my.quantiles)

### Plot quantile predictions from DDM against quantiles from observed data
# Stim 1
par(mfrow=c(1,2), cex=1.2)
plot(data1_q_s1.r1, my.quantiles*observed_prop_correct.s1, pch = 2, 
     ylim=c(0, 1), xlim = c(0.2, 1.0), col="black", type="b",
     ylab = "Cumulative Probability", 
     xlab = "Response Time (sec)", main="Stimulus s1")
lines(data1_q_s1.r2, my.quantiles*(1-observed_prop_correct.s1), pch = 2, 
      col="black", type="b")
lines(pred_q_s1.r1, 
      my.quantiles*pred_prop_correct_DDM.s1, 
      col="red", type = "b")
lines(pred_q_s1.r2, 
      my.quantiles*(1-pred_prop_correct_DDM.s1), 
      col="red", type = "b")
text(x=0.85, y=0.76, "Correct")
text(x=0.8, y=0.3, "Incorrect")
legend("topleft", legend = c("data", "predictions"), pch = c(2, 1), 
       col=c("black", "red"), lty = c(0, 1))

# Stim 2
plot(data1_q_s2.r2, my.quantiles*observed_prop_correct.s2, pch = 2, 
     ylim=c(0, 1), xlim = c(0.2, 1.0), col="black", type="b",
     ylab = "Cumulative Probability", 
     xlab = "Response Time (sec)", main="Stimulus s2")
lines(data1_q_s2.r1, my.quantiles*(1-observed_prop_correct.s2), pch = 2, 
      col="black", type="b")
lines(pred_q_s2.r2, 
      my.quantiles*pred_prop_correct_DDM.s2, 
      col="red", type = "b")
lines(pred_q_s2.r1, 
      my.quantiles*(1-pred_prop_correct_DDM.s2), 
      col="red", type = "b")
text(x=0.85, y=0.81, "Correct")
text(x=0.8, y=0.25, "Incorrect")
legend("topleft", legend = c("data", "predictions"), pch = c(2, 1), 
       col=c("black", "red"), lty = c(0, 1))




############################################################################
####      Parameter estimation on our test dataset,                     ####
####      using RWiener to do MLE.                                      ####
############################################################################

set.seed(1) # reset random seed so results will replicate

# RWiener functions expect the data to be in three-column format:
# q (RT in sec), resp (upper or lower), cond (1, 2, etc.)
q <- data1$RT
resp <- ifelse(data1$R=="r1", "lower", "upper") # define r1 = lower boundary, r2 = upper
cond <- ifelse(data1$S=="s1", 1, 2)
data1.w <- data.frame(q, resp, cond)
table(data1.w$cond, data1.w$resp) # just checking!
#    lower upper
# 1   114    36
# 2    61   289
data1.w$resp <- factor(data1.w$resp)
data1.w$resp <- relevel(data1.w$resp, ref="upper") # upper is reference level
data1.w <- as.wiener(data1.w)  # finally, convert to "data.wiener" object
# subset the data into sublists for the two trial types (conditions)
data1.s1 = subset(data1.w, cond==1)
data1.s2 = subset(data1.w, cond==2)
data1.both = list(data1.s1, data1.s2)

# NB: RWiener's wdm() estimates relative value zr=0..1 as fraction of a
# NB: optimizes $value which is log-likelihood = LLE
# NB: $convergence 1 or 2 indicates convergence, 3=maybe, 4 or 5 fail to converge
# NB: $counts = iterations before convergence 
# NB: optimization method defaults to BFGSD (optim) else nlm else Nelder-Mead (optim)

### Here, we'll consider a simple DDM with h k=5 free parameters: 
# a, Ter, zr, and 2 drift rates, one for s1 and one for s2.  
# NB RWiener returns LLE; this wrapper returns negLLE
# The wrapper function below is adapted from the RWiener documentation.
many_drifts <- function (x, datlist) {
  loss=0
  for (myVar in 1:length(datlist))
    loss = loss - logLik(wdm(datlist[[myVar]], 
                             alpha=x[1], tau=x[2], beta=x[3], delta=x[myVar+3]))
  return(loss)
}

# First, fit the 5-parameter DDM. Here, we will use R's optim()
# function (which uses a Newton-type algorithm) to optimize parameter values.
p_ddm5 <- c(1, 0.1, 0.5, 0, 0) # starting values with two d's
data1_ddm5 <- optim(par=p_ddm5, f=many_drifts, dat=data1.both)
data1_ddm5
# $value 58.7
# $par a 1.174  Ter 0.201  zr 0.603 d.s1 -1.355  d.s2 0.936
# $counts 502
# $convergence 1



# Sanity check: 
# Ter ~200 msec is plausible given observed RTs of 250<RT<1sec.
# Drift rates d.s1<0 (heads toward lower
# boundary) and d.s2>0 (heads toward upper boundary).
# zr ~0.6 means the evidence accumulation process starts ~60%
# of the way from lower to upper boundary, indicating a response bias
# favoring r2; however abs(d.s1)>abs(d.s2) indicates faster evidence
# accumulation for s1 than s2. 


# As an aside, we should get (nearly) identical results using rtdists
# Note, though rtdists returns $convergence=0 if ok, and uses z=z.abs!
rt <- data1.w$q
response <- data1.w$resp
response <- relevel(response, ref="lower") # for rtdists, *lower* is reference level
cond <- data1.w$cond
temp <- data.frame(rt, response, cond)
data1.r <- list(subset(temp, cond==1), subset(temp, cond==2))
head(data1.r)
# define a wrapper calling ddiffusion (instead of deviance.wdm);
# wrapper code adapted from the rtdists documentation.
ll_diffusion <- function(pars, datlist)
{
  loss=0
  for (myVar in 1:length(datlist))
    loss = loss - sum(log(ddiffusion(rt=datlist[[myVar]]$rt, 
                                 response=datlist[[myVar]]$response,
                          a=pars[1], v=pars[3+myVar], t0=pars[2], z=pars[3])))
  return(loss)
}
start <- p_ddm5 # same as RWiener starting vals
names(start) <- c("a", "t0", "z", "v1", "v2") # NB here, z=zabs not zr!
data1_ddm5_2 <- optim(start, f=ll_diffusion, dat=data1.r, method="Nelder-Mead")
data1_ddm5_2
# $par a 1.174  Ter 0.201  zabs 0.708 d.s1 -1.355  d.s2 0.936
# $value 58.7
# $counts 478
# $convergence 0 -- In rtdists, this code means it converged!
# NB zabs=0.708 so zrel = zabs/a = 0.708/1.174=0.603

# for comparison, the results from RWiener:
data1_ddm5
# $par a 1.174  Ter 0.201  zr 0.603 d.s1 -1.355  d.s2 0.935   
# $value 58.7
# $counts 502
# $convergence 1 -- In RWiener, this code means it converged!





# Now we'll go back to RWiener...

# Generate a dataset for parameter recovery check
# cond_list <- c("s1","s2")
parm_recov_s1 <- cbind(rwiener(n=150, alpha=data1_ddm5$par[1],
                               tau=data1_ddm5$par[2],
                               beta=data1_ddm5$par[3],
                               delta=data1_ddm5$par[4]), # v1 for s1
                       cond=1)
parm_recov_s2 <- cbind(rwiener(n=350, alpha=data1_ddm5$par[1],
                               tau=data1_ddm5$par[2],
                               beta=data1_ddm5$par[3],
                               delta=data1_ddm5$par[5]), # v2 for s2
                       cond=2)
parm_recov.w <- as.wiener(rbind(parm_recov_s1, parm_recov_s2))
parm_recov.both <- list(subset(parm_recov.w, cond==1), subset(parm_recov.w, cond==2))

# what have we got?
(x <- table(parm_recov.w$resp, parm_recov.w$cond))
#           s1   s2
# r2=upper  41  294
# r1=lower 109   56   
# i.e., predicted data: 72.7% correct on s1, 84% on s2
# By comparison, original data1 was 76% on s1 and 82.6% on s2

# How do the RTs in the recovered data match the quantiles of observed data?
# How do the quantiles 
my.quantiles <- c(0.1, 0.3, 0.5, 0.7, 0.9) # copied from above
quantiles.dataR <- parm_recov.w %>% group_by(cond, resp) %>% nest() %>%
  mutate(quantiles=map(data, 
                       ~ as.data.frame(t(quantile(.x$q, probs=my.quantiles))))) %>%
  unnest(quantiles) %>%
  gather("quantile", "RT", '10%':'90%') %>%
  arrange(cond, resp)
# quantiles.dataR now contains fields quantile (.1, .3, etc) and RT
# for each combination of stimulus (s1, s2) and response (r1, r2)
#         10%   30%   50%   70%   90% 
#    s1r1 0.314 0.369 0.456 0.585 0.868
#    s1r2 0.257 0.292 0.332 0.436 0.608
#    s2r1 0.287 0.330 0.399 0.539 0.727
#    s2r2 0.265 0.312 0.403 0.536 0.833



# Can the 5-parm DDM recover the generating parameter values?
# Remember that the generating parameters were:
data1_ddm5$par
# $par  1.174  0.201  0.603 -1.355  0.935
# Recover parameters from the predicted data:
parm_recov_ddm5 <- optim(par=p_ddm5, f=many_drifts, dat=parm_recov.both, method="Nelder-Mead")
parm_recov_ddm5
# $value 19   $convergence=1    $counts=502
# $par 1.144  0.204  0.599 -1.146  1.098

# Not bad!




###### FIGURE 6A: Accuracy posterior predictions, DDM using RWiener
par(mfrow=c(1,1))
# Simple plot of accuracy for each stimulus s1, s2
tempX <- list((100*x[2,1])/(x[2,1]+x[1,1]), (100*x[1,2])/(x[1,2]+x[2,2]))
names(tempX) <- c("s1", "s2")
(b <- barplot(unlist(tempX), ylim=c(0,100), ylab="Percent Correct", xlab="Stimulus", main="PREDICTED Accuracy"))
points(x=b, unlist(temp2), col="red")  # overlay dots representing empirical accuracy

###### FIGURE 6B: PDFs for the posterior predictions, from DDM/RWiener
temp_s1r1 <- subset(parm_recov_s1, resp=="lower")
temp_s1r2 <- subset(parm_recov_s1, resp=="upper")
temp_s2r1 <- subset(parm_recov_s2, resp=="lower")
temp_s2r2 <- subset(parm_recov_s2, resp=="upper")
par(mfrow=c(2,2))
# stimulus s1, correct
plot(density(s1r1$RT), main="Stimulus s1: Response r1 (correct)", xlab="RT", ylim=c(0,4))
lines(density(temp_s1r1$q), lwd=2)
legend("topright", c("Data","Model"), lwd=c(1,2), col=c("black", "black"))
# stimulus s2, correct
plot(density(s2r2$RT), main="Stimulus s2: Response r2 (correct)", xlab="RT", ylim=c(0,4))
lines(density(temp_s2r2$q), lwd=2)
legend("topright", c("Data","Model"), lwd=c(1,2), col=c("black", "black"))
# stimulus s1, error
plot(density(s1r2$RT), main="Stimulus s1: Response r2 (error)", xlab="RT", ylim=c(0,4), col="red")
lines(density(temp_s1r2$q), lwd=2, col="red")
legend("topright", c("Data","Model"), lwd=c(1,2), col=c("red", "red"))
# stimulus s2, error
plot(density(s2r1$RT), main="Stimulus s2: Response r1 (error)", xlab="RT", ylim=c(0,4), col="red")
lines(density(temp_s2r1$q), lwd=2, col="red")
legend("topright", c("Data","Model"), lwd=c(1,2), col=c("red", "red"))






############################################################################
####      Parameter estimation on our test dataset,                     ####
####      using DMC for Bayesian parameter estimation.                  ####
############################################################################

set.seed(1) # reset random seed so results will replicate

# load DMC into R workspace - if not already done above
# source ("dmc/dmc.R")

## DEFINE THE DDM MODEL OBJECT AND SOME PRIORS

# load the DDM (only one DMC model can be resident at a time)
load_model("DDM", "ddm.R")


# define task factors (e.g. stimulus classes) & possible responses; 
# NB each response will define one accumulator.
# Map stimuli to correct responses; NB upper/lower boundaries
# will correspond to correct/incorrect
factors <- list(S=c("s1","s2"))               # two possible stimuli are s1 and s2
responses <- c("r1","r2")	                    # two response alternatives are r1 and r1
match.map <- list(M=list(s1="r1",s2="r2"))    # here, r1 is correct s1 and r2 for s2

# define a 5-parameter DDM: a, z, t0 and 2 v (drift rate varies by stimulus)
# Note z=zr (relative value in range 0..1)
p.map.ddm5=list(a="1", v="S", z="1", d="1", sz="1", sv="1", t0="1", st0="1")
constants.ddm5=c(sv=1, sz=0.2, st0=0, d=0)         	       
B.model.ddm5 <- model.dmc(p.map.ddm5, match.map=match.map, constants=constants.ddm5, 
                        factors=factors, responses=responses, type="rd")

# We already have data1 loaded into the workspace
head(data1)
#    S  R    RT corr myType
# 1 s1 r1 0.416    1  s1.r1
# 2 s1 r2 0.261    2  s1.r2
# 3 s1 r1 0.357    1  s1.r1
# 4 s1 r1 0.512    1  s1.r1
# 5 s1 r1 0.390    1  s1.r1
# 6 s1 r1 0.489    1  s1.r1

# create DMC "data.model" object -- same dataset, one for each version of DDM
data1.Bmodel.ddm5 <- data.model.dmc(data1, B.model.ddm5)

# Not run
# As noted above, DMC has some nice plotting routines for data.model 
# objects -- so here is another plot of the observed data
# par(mfrow=c(1,2))
# plot.cell.density(data1.Bmodel.ddm5[data1.Bmodel.ddm5$S=="s1",],C="r1",xlim=c(0,4))
# plot.cell.density(data1.Bmodel.ddm5[data1.Bmodel.ddm5$S=="s2",],C="r2",xlim=c(0,4))

# Define some weakly informative priors: 
# Here, we will use truncated normal distributions for a (mean 1, SD 1, 
# bounded 0..2) and v's (mean 0, SD 2, bounded -5..+5) and uniform distributions
# for z (bounded 0..1) and t0 (bounded 0.1..1.0 sec).
# Note that v>0 is speed to correct boundary (and v<0 is speed to error
# boundary); z is relative response bias (z=0.5 means no bias), and t0
# is lower bound of t0 distribution (if st0=0, Ter=t0, else mean Ter=t0+st0/2)
p.prior.ddm5 <- prior.p.dmc( dists = c("tnorm","tnorm","tnorm","beta","beta"),
                             p1=c(a=1, v.s1=0, v.s2=0, z=1, t0=1), # mean
                             p2=c(a=1, v.s1=2, v.s2=2, z=1, t0=1), # SD
                             lower=c(0, -5, -5, 0, .1), upper=c(2, 5, 5, 1, 1) )


###### FIGURE 7A: DDM priors #######
par(mfcol=c(3,2))
plot.prior("a", p.prior.ddm5, main="a")
plot.prior("v.s2", p.prior.ddm5, main="d.s2")
plot.prior("t0", p.prior.ddm5, main="Ter")
plot.prior("v.s1", p.prior.ddm5, main="d.s1")
plot.prior("z", p.prior.ddm5, main="z")
par(mfcol=c(1,1)) 



### Do sampling

# Generate starting values, run chains -- 1st 200 samples = burn-in with migration on
samples1 <- samples.dmc(nmc=200, p.prior=p.prior.ddm5, data1.Bmodel.ddm5)
samples2 <- run.dmc(samples1, report=50, p.migrate=.05) # NB 3 x k=5 --> 15 chains




###### FIGURE 7B: LLE plot, start of the chains ########
par(mfcol=c(3,2))
plot.dmc(samples2, pll.chain=TRUE, main="Log likelihood estimate (LLE)")  # log-likelihood
###### FIGURE 7B: traceplots for each DDM parameter, start of the chains ######
plot.dmc(samples2)
# Again, for consistency in the manuscript, we will change the 
# variable names "by hand": v-->d and t0-->Ter.



# looks ok, so run for 1000 more with migration off, discarding burn-in
samples3 <- run.dmc(samples.dmc(samples=samples2, nmc=1000), report=50)  



###### FIGURE 7C: LLE plot, end of the chains ########
par(mfcol=c(3,2))
plot.dmc(samples3, pll.chain=TRUE, main="Log likelihood estimate (LLE)")
###### FIGURE 7C: traceplots for each DDM parameter, end of the chains ######
plot.dmc(samples3)        # note the "hairy caterpillars"
# We will change the variable names "by hand" for the ms: v-->d and t0-->Ter.


# Gelman-Rubin R-hat 
gelman.diag.dmc(samples3) 
#        Point est. Upper C.I.
# a          1.02       1.03
# v.s1       1.01       1.02
# v.s2       1.02       1.03
# z          1.02       1.02
# t0         1.02       1.03
# Multivariate psrf 1.03



###### Figure 7D: Posterior distributions 
plot.dmc(samples3, p.prior=p.prior.ddm5, rug.lty=0)   
# To maintain consistency of variable names in the manuscript, we will 
# change the variable names "by hand" to match those in the ms: v-->d and t0-->Ter.





# Report the posterior estimates from DMC
summary.dmc(samples3)    
# Mean a 1.233, v.s1=d.s1 1.490, v.s2=d.s1 1.156, z 0.599, t0=Ter 0.202
# NB: DMC's convention v>0 if slopes toward correct boundary; for ms, 
# we will flip sign, to connote that d.s1<0 slopes to lower boundary = r1.

# Compare these to the values obtained from the MLE approach (RWiener):
data1_ddm5 
# Mean a 1.174, Ter 0.201, z 0.603, d.s1 -1.355, d.s2 0.935
# Generally similar except DMC estimates somewhat larger (more
# extreme) drift rates
# Cautionary note: generally, drift rates should be considered relative 
# rather than absolute values (i.e. compare d.s1 vs. d.s2... or compare 
# across subjects under same model), since
# different packages/approaches may use different scaling factors!




###### TABLE 1: text only ######

###### TABLE 2: text only ######

 

###### TABLE 3: Report parameter estimates from DMC; compare RWiener/rtdists #########
x <- summary.dmc(samples3) 
dmc_ests <- x$statistics[,1]
names(dmc_ests) <- c("a", "d.s1", "d.s2", "z", "Ter")
# flip d.s1 so slope is negative
dmc_ests[2] <- -1 * dmc_ests[2]
dmc_ests
#     a   d.s1  d.s2     z   Ter 
# 1.233 -1.490 1.156 0.599 0.202 
RW_ests <- data1_ddm5$par
names(RW_ests) <- c("a","Ter","z","d.s1","d.s2")
RW_ests <- RW_ests[c("a","d.s1","d.s2","z","Ter")]
RW_ests
#     a   d.s1   d.s2      z    Ter 
# 1.174 -1.355  0.935  0.603  0.201
true_vals <- c(1.2,-1.25,1.1,0.6,0.2)
names(true_vals) <- names(RW_ests)

table3 <- as.data.frame(t(data.frame(RW_ests, dmc_ests, true_vals)))
table3$myLabels <- c("MLE (using RWiener)", "Bayesian (using DMC)", "True (generating) values")

myTable3 <- table3 %>%
  gt(rowname_col="myLabels") %>%
  tab_header(title="DDM: Estimated Parameter Values")
myTable3
myTable.filename <- paste(data.pathname, "DDM_review_Table3.rtf", sep="")
gtsave(myTable3, myTable.filename)   # or copy image from RStudio plot window








###########################################################################
####  Multiple files: parameter estimation & model validation.         ####  
###########################################################################


set.seed(1) # re-set random seed so results will replicate

# NB files empdataX.csv (X=1..10) each have: one row per trial, 
# column cond (s1/s2), resp (lower/upper), rt (in sec) -- already in
# format to analyze via rtdists!

# First, use 5-parm DDM to est pararameters for each empirical datafile
myNumSubj <- 10
myNparms <- 5
empirical_ests <- data.frame(matrix(NA, nrow=myNumSubj, ncol=myNparms+3))
names(empirical_ests) <- c("subjID", "a", "z", "d.s1", "d.s2", "Ter", "converge", "LLE")
start <- c(1.0, 0.1, 0.5, 0.0, 0.0)  # We'll use same starting point as defined above
names(start) <- c("a", "t0", "z", "v1", "v2") # NB here, z=zabs not zr!
for (i in 1:myNumSubj) {
  # read in one empirical datafile
  infilename <- paste(data.pathname, "empdata", i, ".csv", sep="")
  emp.data <- read.csv(infilename, header=TRUE, sep=",")
  # estimate parms, using the ll_diffusion wrapper defined above (calls rtdist's ddiffusion)
  data.to.model <- list(subset(emp.data, cond=="s1"), subset(emp.data, cond=="s2"))
  one_est <- optim(start, f=ll_diffusion, dat=data.to.model, method="BFGS")
  # save to empirical_ests
  parms <- one_est$par
  empirical_ests[i,"subjID"] <- i
  empirical_ests[i,"a"] <- parms["a"]; 
  empirical_ests[i,"Ter"] <- parms["t0"];
  empirical_ests[i,"z"] <- parms["z"]/parms["a"]; # rtdists returns z.abs; convert to z.rel
  empirical_ests[i, "d.s1"] <- parms["v1"];
  empirical_ests[i,"d.s2"] <- parms["v2"]; 
  empirical_ests[i,"converge"] <- one_est$convergence
  empirical_ests[i,"LLE"] <- -1*one_est$value  #rtdists returns negLLE; convert to LLE
}
(empirical_ests <- round(empirical_ests, digits=2))
#    subjID    a    z  d.s1 d.s2  Ter converge LLE
# 1       1 1.09 0.46 -3.93 1.27 0.21        0 185
# 2       2 1.91 0.48 -3.66 1.97 0.25        0  70
# 3       3 1.85 0.51 -5.30 1.80 0.16        0 129
# 4       4 1.19 0.55 -3.62 2.09 0.17        0 287
# 5       5 1.10 0.46 -4.01 1.27 0.21        0 190
# 6       6 1.65 0.49 -2.41 2.43 0.25        0 140
# 7       7 1.20 0.44 -2.01 2.02 0.21        0 129
# 8       8 0.78 0.40 -1.01 1.00 0.23        0 220
# 9       9 0.83 0.53 -3.20 1.53 0.18        0 332
# 10     10 0.85 0.43 -1.20 0.11 0.18        0 136
# Note: remember, using rtdist's ll_diffusion(), convergence=0 means ok.


###### TABLE 4: Hypothetical study, DDM results ######
# Compute means and SD 
myMeans <- c(NA, mean(empirical_ests$a), mean(empirical_ests$z), 
             mean(empirical_ests$d.s1), 
             mean(empirical_ests$d.s2), mean(empirical_ests$Ter),
             NA, mean(empirical_ests$LLE))
myMeans <- round(myMeans, digits=2)
mySD <- c(NA, sd(empirical_ests$a), sd(empirical_ests$z), sd(empirical_ests$d.s1),
          sd(empirical_ests$d.s2), sd(empirical_ests$Ter), 
          NA, sd(empirical_ests$LLE))   
mySD <- round(mySD, digits=2)
# Create a master table, including Mean/SD
myEstParm <- rbind(empirical_ests, myMeans, mySD)
myEstParm$subjID[11] <- "Mean"
myEstParm$subjID[12] <- "SD"
myTable4 <-
  myEstParm[,!(names(myEstParm) %in% c("converge"))] %>%
  gt(rowname_col="subjID") %>%
  cols_align(align="center") %>%
  tab_spanner("Estimated Parameters", columns=c("a","z","d.s1","d.s2","Ter")) %>%
  cols_label(subjID="Subject ID") 
myTable4
myTable.filename <- paste(data.pathname, "DDM_review_Table4.rtf", sep="")
gtsave(myTable4, myTable.filename)   # or copy image from RStudio plot window



# Predictive check
# We then use 5-parm DDM to generate one simulated dataset per each set of values
for (i in 1:myNumSubj) { # for each subject ID
  myVals <- empirical_ests[empirical_ests$subjID==i,]
  # 150 s1 trials (using d.s1)... note rtdists uses z.abs, not z.rel, so convert to z.abs!
  temp.s1 <- rdiffusion(n=150, a=myVals$a, v= myVals$d.s1, t0=myVals$Ter, z=myVals$z*myVals$a)
  # 350 s2 trials (using d.s2)
  temp.s2 <- rdiffusion(n=350, a=myVals$a, v=myVals$d.s2, t0=myVals$Ter, z=myVals$z*myVals$a)
  # concatenate into an "experiment" of 500 trials
  temp.s1$cond="s1"
  temp.s2$cond="s2"
  temp.all=rbind(temp.s1, temp.s2) 
  # save as simulated dataset; note column names cond=S, response=R, rt=RT;
  # we won't change these column labels, since they match what rtdists will expect, below!
  outfilename <-paste(data.pathname, "simdata", i, ".csv", sep="")
  write.csv(subset(temp.all, select=c(cond, response, rt)), file=outfilename, row.names=FALSE)
}


# Posterior predictive check: do simulated data look like empirical data?
# Read in each empirical datafile; calc accuracy and median RTs; 
# then do same for simulated; then plot (similar to Figs 5 & 9)
set.seed(0)
myColNames <- c("s1.accuracy", "s2.accuracy", "s1.correctRT", 
                           "s1.incorrectRT", "s2.correctRT", "s2.incorrectRT")
myEmpirical <- matrix(nrow=10, ncol=6, dimnames=list(c(1:10), myColNames))
for (i in 1:10) {
  infilename <- paste(data.pathname, "empdata", i, ".csv", sep="")
  emp.data <- read.csv(infilename, header=TRUE, sep=",")
  myS1 <- subset(emp.data, cond=="s1")
  myS2 <- subset(emp.data, cond=="s2")
  myS1.correct <- subset(myS1, response=="lower")
  myS1.incorrect <- subset(myS1, response=="upper")
  myS2.correct <- subset(myS2, response=="upper")
  myS2.incorrect <- subset(myS2, response=="lower")
  myEmpirical[i,1] = 100*length(myS1.correct$rt)/length(myS1$rt)
  myEmpirical[i,2] = 100*length(myS2.correct$rt)/length(myS2$rt)
  if (is.na(length(myS1.correct))) myEmpirical[i,3] = NA else myEmpirical[i,3]=median(myS1.correct$rt)
  if (is.na(length(myS1.incorrect))) myEmpirical[i,4] = NA else myEmpirical[i,4]=median(myS1.incorrect$rt)
  if (is.na(length(myS2.correct))) myEmpirical[i,5] = NA else myEmpirical[i,5]=median(myS2.correct$rt)
  if (is.na(length(myS2.incorrect))) myEmpirical[i,6] = NA else myEmpirical[i,6]=median(myS2.incorrect$rt)
}
mySimulated <- matrix(nrow=10, ncol=6, dimnames=list(c(1:10), myColNames))
for (i in 1:10) {
  infilename <- paste(data.pathname, "simdata", i, ".csv", sep="")
  sim.data <- read.csv(infilename, header=TRUE, sep=",")
  myS1 <- subset(sim.data, cond=="s1")
  myS2 <- subset(sim.data, cond=="s2")
  myS1.correct <- subset(myS1, response=="lower")
  myS1.incorrect <- subset(myS1, response=="upper")
  myS2.correct <- subset(myS2, response=="upper")
  myS2.incorrect <- subset(myS2, response=="lower")
  mySimulated[i,1] = 100*length(myS1.correct$rt)/length(myS1$rt)
  mySimulated[i,2] = 100*length(myS2.correct$rt)/length(myS2$rt)
  if (is.na(length(myS1.correct))) mySimulated[i,3] = NA else mySimulated[i,3]=median(myS1.correct$rt)
  if (is.na(length(myS1.incorrect))) mySimulated[i,4] = NA else mySimulated[i,4]=median(myS1.incorrect$rt)
  if (is.na(length(myS2.correct))) mySimulated[i,5] = NA else mySimulated[i,5]=median(myS2.correct$rt)
  if (is.na(length(myS2.incorrect))) mySimulated[i,6] = NA else mySimulated[i,6]=median(myS2.incorrect$rt)
}
myEmpirical <- as.data.frame(myEmpirical)
mySimulated <- as.data.frame(mySimulated)

###### FIGURE 8: Scatterplots of Empirical data vs. Simulated data #####
par(mfrow=c(1,3))
plot(myEmpirical$s1.accuracy,mySimulated$s1.accuracy, col="black", xlim=c(50,100), ylim=c(50,100),
     xlab="Empirical data", ylab="Simulated data", main="Accuracy (% Correct)")
points(myEmpirical$s2.accuracy,mySimulated$s2.accuracy, col="red")
legend(x=50, y=100, legend=c("s1", "s2"), pch=c(1,1), col=c("black", "red"))
abline(coef=c(0,1))
plot(myEmpirical$s1.correct,mySimulated$s1.correct, col="black", xlim=c(0,1), ylim=c(0,1),
     xlab="Empirical data", ylab="Simulated data", main="Median RT (sec) for correct responses")
points(myEmpirical$s2.correct,mySimulated$s2.correct, col="red")
legend(x=0, y=1, legend=c("s1", "s2"), pch=c(1,1), col=c("black", "red"))
abline(coef=c(0,1))
plot(myEmpirical$s1.incorrect,mySimulated$s1.incorrect, col="black", xlim=c(0,1), ylim=c(0,1),
     xlab="Empirical data", ylab="Simulated data", main="Median RT (sec) for incorrect responses")
points(myEmpirical$s2.incorrect,mySimulated$s2.incorrect, col="red")
legend(x=0, y=1, legend=c("s1", "s2"), pch=c(1,1), col=c("black", "red"))
abline(coef=c(0,1))
cor(myEmpirical$s1.accuracy, mySimulated$s1.accuracy)
cor(myEmpirical$s2.accuracy, mySimulated$s2.accuracy)
cor(myEmpirical$s1.correct, mySimulated$s1.correct)
cor(myEmpirical$s1.incorrect, mySimulated$s1.incorrect, use="complete.obs")
cor(myEmpirical$s2.correct, mySimulated$s2.correct)
cor(myEmpirical$s2.incorrect, mySimulated$s2.incorrect, use="complete.obs")


# Now, fit the 5-parm DDM to each simulated datafile, and save the parm estimates
recovery_ests <- data.frame(matrix(NA, nrow=myNumSubj, ncol=myNparms+3))
names(recovery_ests) <- c("subjID","a", "z", "d.s1", "d.s2", "Ter", "converge", "LLE")
start <- c(1.0, 0.1, 0.5, 0.0, 0.0)  # We'll use same starting point as defined above
names(start) <- c("a", "t0", "z", "v1", "v2") # NB here, z=zabs not zr!
for (i in 1:myNumSubj) {
  # read in one simulated datafile
  infilename <- paste(data.pathname, "simdata", i, ".csv", sep="")
  sim.data <- read.csv(infilename, header=TRUE, sep=",")
  # estimate parms, using the ll_diffusino wrapper defined above (calls rtdist's ddiffusion)
  data.to.model <- list(subset(sim.data, cond=="s1"), subset(sim.data, cond=="s2"))
  one_est <- optim(start, f=ll_diffusion, dat=data.to.model, method="BFGS")
  # save to recovery_ests
  parms <- one_est$par
  recovery_ests[i,"subjID"] <- i
  recovery_ests[i,"a"] <- parms["a"]; 
  recovery_ests[i,"Ter"] <- parms["t0"];
  recovery_ests[i,"z"] <- parms["z"]/parms["a"]; # rtdists returns z.abs; convert to z.rel
  recovery_ests[i, "d.s1"] <- parms["v1"]; 
  recovery_ests[i,"d.s2"] <- parms["v2"]; 
  recovery_ests[i,"converge"] <- one_est$convergence
  recovery_ests[i,"LLE"] <- -1*one_est$value
}
(recovery_ests <- round(recovery_ests, digits=2))
#    subjID    a    z  d.s1 d.s2  Ter converge   LLE
# 1       1 1.06 0.48 -4.34 1.20 0.21        0 198.8
# 2       2 1.89 0.50 -3.91 1.80 0.25        0  53.2
# 3       3 1.74 0.52 -5.14 1.73 0.17        0 129.0
# 4       4 1.23 0.55 -3.69 1.97 0.16        0 264.5
# 5       5 1.09 0.47 -3.69 1.27 0.21        0 156.8
# 6       6 1.62 0.49 -2.31 2.50 0.25        0 144.8
# 7       7 1.22 0.43 -1.82 1.78 0.21        0  57.6
# 8       8 0.80 0.41 -0.83 0.64 0.23        0 180.3
# 9       9 0.82 0.53 -2.87 1.51 0.18        0 334.1
# 10     10 0.85 0.41 -1.07 0.27 0.18        0 127.0
# Note: remember, using rtdist's ll_diffusion(), convergence=0 means ok.


###### TABLE 5: Parameter recovery study results ######
# Compute means and SD 
myMeans <- c(NA, mean(recovery_ests$a), mean(recovery_ests$z), 
             mean(recovery_ests$d.s1), 
             mean(recovery_ests$d.s2), mean(recovery_ests$Ter),
             NA, mean(recovery_ests$LLE))
myMeans <- round(myMeans, digits=2)
mySD <- c(NA, sd(recovery_ests$a), sd(recovery_ests$z), sd(recovery_ests$d.s1),
          sd(recovery_ests$d.s2), sd(recovery_ests$Ter), 
          NA, sd(recovery_ests$LLE))   
mySD <- round(mySD, digits=2)
# Create a master table, including Mean/SD
myRecovParm <- rbind(recovery_ests, myMeans, mySD)
myRecovParm$subjID[11] <- "Mean"
myRecovParm$subjID[12] <- "SD"
myTable5 <-
  myRecovParm[,!(names(myRecovParm) %in% c("converge"))] %>%
  gt(rowname_col="subjID") %>%
  cols_align(align="center") %>%
  tab_spanner("Recovered Parameters", columns=c("a","z","d.s1","d.s2","Ter")) %>%
  cols_label(subjID="Subject ID") 
myTable5
myTable.filename <- paste(data.pathname, "DDM_review_Table5.rtf", sep="")
gtsave(myTable5, myTable.filename)   # or copy image from RStudio plot window




###### FIGURE 9: Scatterplot estimated vs. recovered parameters
r_a <- round(cor(recovery_ests$a, empirical_ests$a), digits=2)
r_z <- round(cor(recovery_ests$z, empirical_ests$z), digits=2)
r_d.s1 <- round(cor(recovery_ests$d.s1, empirical_ests$d.s1), digits=2)
r_d.s2 <- round(cor(recovery_ests$d.s2, empirical_ests$d.s2), digits=2)
r_Ter <- round(cor(recovery_ests$Ter, empirical_ests$Ter), digits=2)
par(mfrow=c(3,2))
xlab="Estimated Parameters"; ylab="Recovered Parameters"
plot(recovery_ests$a, empirical_ests$a, 
     main=paste("a: r=",r_a), xlab=xlab, ylab=ylab)
plot(recovery_ests$d.s1, empirical_ests$d.s1, 
     main=paste("d.s1: r=",r_d.s1), xlab=xlab, ylab=ylab)
plot(recovery_ests$d.s2, empirical_ests$d.s2, 
     main=paste("d.s2: r=",r_d.s2), xlab=xlab, ylab=ylab)
plot(recovery_ests$z, empirical_ests$z, 
     main=paste("z: r=",r_z), xlab=xlab, ylab=ylab)
plot(jitter(recovery_ests$Ter), jitter(empirical_ests$Ter), 
     main=paste("Ter: r=",r_Ter), xlab=xlab, ylab=ylab)







############################################################################
####      Approaches for model selection                                ####
############################################################################


set.seed(1) # re-set random seed so results will replicate


# We already used rtdists to estimate parameter values for our 10
# subjects using DDM-5 (5 free parameters: a, z, Ter, and 2 drift rates)
(DDM5_ests <- empirical_ests)


# Now, we'll consider a simpler DDM with k=4 free parameters: 
# a, Ter, zr, and 1 drift rate d... but using -d for s1 and +d for s2.  
# NB using rtdists; again, we need to convert to/from z.abs, and
# are returning negLLE = negative of logLik
# Wrapper below adapted from rtdists documentation.
single_drift <- function (pars, datalist) {
  loss = 0 
  loss = loss - sum(log(ddiffusion(rt=datalist[[1]]$rt, 
                                   response=datalist[[1]]$response,
                                   a=pars[1], v= -1*pars[4], t0=pars[2], z=pars[3]*pars[1])))
  loss = loss - sum(log(ddiffusion(rt=datalist[[2]]$rt, 
                                   response=datalist[[2]]$response,
                                   a=pars[1], v=pars[4], t0=pars[2], z=pars[3]*pars[1])))
  return(loss)
}


# Now, use 4-parm DDM to est parameters for each empirical datafile
myNumSubj <- 10
myNparms <- 4
DDM4_ests <- data.frame(matrix(NA, nrow=myNumSubj, ncol=myNparms+4))
names(DDM4_ests) <- c("subjID", "a", "z", "d.s1", "d.s2", "Ter", "converge", "LLE")
start4 <- c(1.0, 0.1, 0.5, 0)  # New starting points 
names(start4) <- c("a", "t0", "z", "v") # NB here, z=zabs not zr!
for (i in 1:myNumSubj) {
  # read in one empirical datafile
  infilename <- paste(data.pathname, "empdata", i, ".csv", sep="")
  emp.data <- read.csv(infilename, header=TRUE, sep=",")
  # estimate parms, using the ll_diffusion wrapper defined above (calls rtdist's ddiffusion)
  data.to.model <- list(subset(emp.data, cond=="s1"), subset(emp.data, cond=="s2"))
  one_est <- optim(start4, f=single_drift, dat=data.to.model, method="BFGS")
  # save to empirical_ests
  parms <- one_est$par
  DDM4_ests[i,"subjID"] <- i
  DDM4_ests[i,"a"] <- parms["a"]; 
  DDM4_ests[i,"Ter"] <- parms["t0"];
  DDM4_ests[i,"d.s1"] <- -1*parms["v"]; DDM4_ests[i,"d.s2"] <- parms["v"]
  DDM4_ests[i, "z"] <- parms["z"]/parms["a"];
  DDM4_ests[i, "converge"] <- one_est$convergence
  DDM4_ests[i,"LLE"] <- -1*one_est$value  #rtdists returns negLLE; convert to LLE
}
(DDM4_ests <- round(DDM4_ests, digits=2))
#    subjID    a    z  d.s1 d.s2  Ter converge   LLE
# 1       1 1.08 0.36 -1.87 1.87 0.22        0 150.6
# 2       2 1.88 0.22 -2.35 2.35 0.24        0  49.9
# 3       3 1.74 0.22 -2.36 2.36 0.15        0  69.2
# 4       4 1.17 0.43 -2.51 2.51 0.17        0 273.6
# 5       5 1.08 0.36 -1.88 1.88 0.22        0 153.0
# 6       6 1.65 0.30 -2.42 2.42 0.25        0 140.5
# 7       7 1.20 0.36 -2.02 2.02 0.21        0 129.5
# 8       8 0.78 0.51 -1.00 1.00 0.23        0 220.4
# 9       9 0.83 0.59 -2.01 2.01 0.18        0 319.4
# 10     10 0.85 0.47 -0.45 0.45 0.18        0 128.9
# Again, remember rtdist's ddiffusion converge=0 if ok!



###### TABLE 6: Model comparison: DDM5 vs. DDM4 ######
# Compute means and SD 
myMeans <- c(NA, mean(DDM4_ests$a), mean(DDM4_ests$z), 
             mean(DDM4_ests$d.s1), 
             mean(DDM4_ests$d.s2), mean(DDM4_ests$Ter),
             NA, mean(DDM4_ests$LLE))
myMeans <- round(myMeans, digits=2)
mySD <- c(NA, sd(DDM4_ests$a), sd(DDM4_ests$z), sd(DDM4_ests$d.s1),
          sd(DDM4_ests$d.s2), sd(DDM4_ests$Ter), 
          NA, sd(DDM4_ests$LLE))   
mySD <- round(mySD, digits=2)
# Create a master table, including Mean/SD
myDDM4 <- rbind(DDM4_ests, myMeans, mySD)
myDDM4$subjID[11] <- "Mean"
myDDM4$subjID[12] <- "SD"
# add DDM5 LLE, for comparison
myDDM4$LLEfromDDM5 <- myEstParm$LLE

myTable6 <-
  myDDM4[,!(names(myDDM4) %in% c("converge"))] %>%
  gt(rowname_col="subjID") %>%
  cols_align(align="center") %>%
  tab_spanner("Estimated Parameters, using DDM-4", columns=c("a","z","d.s1","d.s2","Ter")) %>%
  cols_label(subjID="Subject ID", LLE="LLE (from DDM-4)", LLEfromDDM5="LLE (from DDM-5)") 
myTable6
myTable.filename <- paste(data.pathname, "DDM_review_Table6.rtf", sep="")
gtsave(myTable6, myTable.filename)   # or copy image from RStudio plot window




# The larger model fits better (larger LLE on average) but at cost of
# more complexity (more free parameters)

# One way to penalize for complexity is Akaike Information Criterion (AIC)
# AIC = 2k-2LLE
(AIC.ddm5 <- 2*5-2*DDM5_ests$LLE)   
(AIC.ddm4 <- 2*4-2*DDM4_ests$LLE)      
AIC.ddm5 - AIC.ddm4  # smaller (more negative) numbers --> DDM-5 is better 

(meanDDM5_LLE <- round(mean(DDM5_ests$LLE), digits=2))   # 167
(meanDDM4_LLE <- round(mean(DDM4_ests$LLE), digits=2))   # 164
(meanDDM5_AIC <- round(mean(AIC.ddm5), digits=2))        # -354 smaller!
(meanDDM4_AIC <- round(mean(AIC.ddm4), digits=2))        # -319
# AIC is somewhat smaller (more negative) for larger model -- i.e. the added complexity is "worth it."


# Some prefer Bayesian Information Criterion (BIC) which also adjusts
# for number of trials (which influences absolute value of LLE)
# Formally, BIC provides estimate of the posterior probability being 
# true (lower BIC -> more likely that this is "true" model). BIC is only
# valid if n>>k (number of observations, here trials >> number of free
# parameters. As a rule (Kass & Raftery, 1995), if difference between BIC
# between a complex and more simple models is:
#   <2:   more complex model is "not worth it"
#   2-6:  positive evidence in favor of more complex model
#   6-10: strong evidence in favor of more complex model
#   10+:  very strong evidence in favor of more complex model

# BIC = k*log(n)-2LLE 
n <- 500   # assuming n=500 for all subject/datafiles, log(n)=6.21

(BIC.ddm5 <- 5*log(n)-2*DDM5_ests$LLE)   
(BIC.ddm4 <- 4*log(n)-2*DDM4_ests$LLE)      
BIC.ddm5 - BIC.ddm4  # smaller (more negative) numbers --> DDM-5 is better 

(meanDDM5_BIC <- round(mean(BIC.ddm5), digits=2))    # -333 -- smaller!
(meanDDM4_BIC <- round(mean(BIC.ddm4), digits=2))    # -302
# BIC is smaller (more negative) for larger model -- i.e. added complexity is "worth it."



###### TABLE 7: AIC AND BIC - not included in MS ######
myLabels <- c("DDM-5 (5 parameters including 2 drift rates)",
              "DDM-4 (4 parameters including 1 drift rate)")
myMeanLLE <- c(meanDDM5_LLE, meanDDM4_LLE)
myK <- c(5,4)
myMeanAIC <- c(meanDDM5_AIC, meanDDM4_AIC)
myN <- c(500,500)
myMeanBIC <- c(meanDDM5_BIC, meanDDM4_BIC)
table7 <- data.frame(myLabels, myMeanLLE, myK, myMeanAIC, myN, myMeanBIC)
myTable7 <-
  table7 %>%
  gt(rowname_col="myLabels") %>%
  cols_align(align="center") %>%
  cols_label(myLabels="Model", myMeanLLE="Mean LLE", myK="k (# parameters)",
             myMeanAIC="Mean AIC", myN="n (#trials)", myMeanBIC="Mean BIC")
myTable7
myTable.filename <- paste(data.pathname, "DDM_review_Table7.rtf", sep="")
gtsave(myTable7, myTable.filename)   # or copy image from RStudio plot window




###### FIGURE 10: REPORT MODEL PARAMETERS ######
# for plotting purposes, we'll plot d.s1 as >0
temp2 <- empirical_ests
temp2$d.s1 <- -1*temp2$d.s1
temp2$Group <- c(rep("subjects 1-5", 5), rep("subjects 6-10", 5))
temp.ad <- melt(temp2, id.vars=c("subjID", "Group"),
             measure.vars=c("a", "d.s1", "d.s2"),
             variable.name="Parameter")
temp.zTer <- melt(temp2, id.vars=c("subjID", "Group"),
                measure.vars=c("z", "Ter"),
                variable.name="Parameter")

# Boxplot of all 5 parameter estimates from DDM-5 on our n=10 subjects
pd<-position_dodge(1)
g.ad <- ggplot(temp.ad, aes(x=Parameter, y=value, fill=Group)) +
  geom_boxplot(position=pd, aes(fill=Group)) +
  scale_fill_manual(values=c("#2171b5","white"))  +
  xlab("Parameter") +
  ylab("Median Value") +
  theme_classic() + theme(legend.position="none")

# Plot z and Ter separately, as they are on different scales than a, d
g.zTer <- ggplot(temp.zTer, aes(x=Parameter, y=value, fill=Group)) +
  geom_boxplot(position=pd, aes(fill=Group)) +
  scale_fill_manual(values=c("#2171b5","white"))  +
  xlab("Parameter") +
  ylab("Median Value") +
  theme_classic() + theme(legend.position=c(0.8,0.8))
# ... or, plot separately, given that there are different scales

grid.arrange(g.ad, g.zTer, ncol=2, nrow=1)




############################################################################
####      The end.                                                      ####
############################################################################
```
